---
title: "CHOE.JONGWOOK-HW1"
author: "JONG-WOOK-CHOE"
output:
  pdf_document: default
  html_document: default
---
BOOK HOMEWORK

1. a) Grade(A, B, C, D, E, F) 
      employment status(Unemployed, Employed)
      AU, BU, CU, DU, FU, AE, BE, CE, DE, FE
      
   b) Cancer status(has Cancer, not Cancer)
      age group(Young, Middle, Old)
      hCY, hCM, hCO, nCY, nCM, nCO
      
   c) Smoking status(smoker, non-smoker)
      drug activity(Recently used, used in Past, Never used)
      sR, sP, sN, nR, nP, nN
      
   d) Intelligence of dog(high, medium, low)
      type of dog(Border collie, German shepard, Dachshund)
      hB, hG, hD, mB, mG, mD, lB, lG, lD
   
2. a) mean = 4*4+3 = 19
      standard deviation = 8*4 = 32
      
   b) mean = 2*4-10 = -2
      standard deviation = 8*2 = 16
      
   c) mean = -4+1/4 = -3.75
      standard deviation = 8
      
   d) mean = -(1/4)*4+1/4 = -0.75
      standard deviation = 8*(1/4) = 2
   
3. a) mean = 20
      variance = 25/10 = 2.5
      
   b) mean = 20*10 = 200
      variance = 250
      
   c) mean = 20b+a
      variance = 2.5$b^2$
      
   d) mean = -395
      variance = 1000
   
4. a) mean = $\mu_1-\mu_2$
      standard deviation = $\frac{\sqrt{\sigma_1^2+\sigma_2^2}}{10}$

   b) mean = $\mu_1+\mu_2$
      standard deviation = $\frac{\sqrt{\sigma_1^2+\sigma_2^2}}{10}$
      
   c) mean = $\frac{\mu_1+\mu_2}{2}-\mu_3$
      standard deviation = $\frac{\sqrt{\sigma_1^2+\sigma_2^2-4\sigma_3^2}}{20}$  
      
   d) mean = $\mu_1+\mu_3-\mu_2$ 
      standard deviation = $\frac{\sqrt{\sigma_1^2+\sigma_3^2-4\sigma_2^2}}{10}$
   
5. a) $H_0$ is when Average systolic blood pressure differ by smoking status
      $H_a$ is when Average systolic blood pressure doesn't differ by smoking status
      $H_0:\mu_1=\mu_2$
      $H_a:\mu_1\not=\mu_2$
      when $\mu_1$ is mean systoloc blood pressure of smoker
      when $\mu_2$ is mean systoloc blood pressure of non smoker
      
   b) t-test = $\frac{\bar{Y_1}-\bar{Y_2}}{\sigma_{(\bar{Y_1}-\bar{Y_2})}}=\frac{\bar{Y_1}-\bar{Y_2}}{\sqrt{Sp^2(\frac{1}{n_1}+\frac{1}{n_2})}}$
             = $\frac{150.03-130.18}{\sqrt{6.070465}}$
             = 4.403710323
             
   c) p-value = $2P(T>4.403710323)=0.000013$ < 0.0005
      From the t-table the number is large so it is less then the closest t-statistic value.
      
   d) When $\alpha=0.05$ p-value is less than alpha therefore reject $H_0$, null hypothesis, and support $H_a$ , alternative hypothesis. 
      
6. a) Type1 error arises if it conclude to reject $H_0$ when $H_0$ is true.
      
   b) Type2 error arises if it conclude not to reject $H_0$ when $H_0$ is wrong.
   
   c) 99% confidence interval for the true difference in average systolic blood pressure is (4.451,17.249)
   
   d) Since 99% confidence interval is (4.451,17.249) and doesn't contain zero, it means that the average systolic blood pressure of smokers and no smokers are unequal. Also it shows the the average systolic blood pressure of smokers is higher than that of non-smokers.
   
   e) 17.249 is the largest difference between the two groups. It means that at most 17.249 the average systolic blood pressures of smokers be higher than that of non smokers.
   
7. a) This study experimental because students are randomly assigned to two different types of instruction.

   b) The primarey variable, response variable, is the standardized mathematics test at the end of the term.
   
   c) There are three explanatory variables. First is the high school with three factor levels, types of instructions assigned with two factor levels, and sections in each schools with four factor levels.
   
   d) 
|           |section_a    |section_b    |section_c    |section_d    | 
|           |S     |C     |S     |C     |S     |C     |S     |C     | 
|:----------|------------:|------------:|------------:|------------:|
|School_1   |1aS   |1aC   |1bS   |1bC   |1cS   |1cC   |1dS   |1dC   |      
|School_2   |2aS   |2aC   |2bS   |2bC   |2cS   |2cC   |2dS   |2dC   | 
|School_3   |3aS   |3aC   |3bS   |3bC   |3cS   |3cC   |3dS   |3dC   | 
      
      when section (a, b, c), type of instruction (Sandard, Computer-based), and school (1, 2, 3)
      1aS, 1aC, 1bS, 1bC, 1cS, 1cC, 1dS, 1dC, 2aS, 2aC, 2bS, 2bC, 2cS, 2cC, 2dS, 2dC, 3aS, 3aC, 3bS, 3bC, 3cS, 3cC, 3dS, 3dC.    

   
8. a) This study is observational because the researchers observes and examined the relationship between variables without influence or intrusion.

   b) The primary variable, response variable, is the time required in physical therapy until successful rehabilitation.
   
   c) There are only one explanatory variables which is the prior physical fitness status with three factor levels.
   
   d) bA, 0A, aA (below Average, Average, above Average)


R HOMEWORK

I

```{r, echo=FALSE}
newdata = read.csv("/Users/Bosco/Desktop/WINTER2021/STA 106/data/GSK.csv")
library(MASS)
summary(newdata)
a.means = aggregate(sbp~stress,data=newdata, mean)
a.means
print("(a) Group that has the highest average is high stress level.")
b.sds = aggregate(sbp~stress,data=newdata, sd)
b.sds
print("(b) It appears the standard deviations are approximately equal.")
c.means = aggregate(age~exercise,data=newdata, mean)
c.means
print("(c) Group that has the lowest average age is is medium exercise group.")
d.sds = aggregate(age~exercise,data=newdata, sd)
d.sds
print("(d) It appears the standard deviations of age by exercise level differ the most from the group L, low exercise level")
```

II

```{r, echo=FALSE}
boxplot(sbp~educatn, data=newdata, main="Systolic Blood Pressure by Education",horizontal = TRUE)
print("(a) There is a trend that all groups are nearing each others")
library(ggplot2)
ggplot(newdata, aes(x=sbp))+geom_histogram(binwidth=2,color="black",fill="white")+facet_grid(married~.)+ggtitle("Systolic Blood Pressure by Marriage")
print("(b) Both group vary more than the other")
e_s=table(newdata$exercise,newdata$stress)
mosaicplot(e_s,main="Exercise vs Stress")
print("(c) The highest proportion of low stress is in the low exercise group")
m_s=table(newdata$married,newdata$stress)
mosaicplot(m_s,main="Marriage vs Stress")
print("(d) The highest proportion of low stress is in the married group")
```

III

```{r, echo=FALSE}
table(newdata$exercise)
print("(a) Low exercise group had the most subject")
table(newdata$stress,newdata$exercise)
print("(b) Medium stress group had the most highly educated subjects")
s_a = aggregate(age~stress,data=newdata, mean)
s_a
print("(c) Medium stress group had the highest average age")
g_a = aggregate(sbp~gender,data=newdata, mean)
g_a
print("(d) Female had the lowest average systolic blood pressure")
```

IV

```{r,echo=FALSE}
result=t.test(sbp~married,data=newdata,var.equal=TRUE,level=0.95)
ts=round(result$statistic,4)
pval=result$p.value
CI=round(result$conf.int,4)
ts
print("(a) -1.3554")
pval
print("(b) 0.1758972")
CI
print("(c) (-8.3146  1.5259)")
print("(d) The confidence interval includes 0 meaning there are no significant difference between the groups")
print("(e) Since p value is larger than alpha null hypothesis is accepted and meaning that there are no relationship between two groups")
```

## Appendix

```{r,eval=FALSE}
newdata = read.csv("/Users/Bosco/Desktop/WINTER2021/STA 106/data/GSK.csv")
library(MASS)
summary(newdata)
a.means = aggregate(sbp~stress,data=newdata, mean)
print("(a) Group that has the highest average is high stress level.")
b.sds = aggregate(sbp~stress,data=newdata, sd)
print("(b) It appears the standard deviations are approximately equal.")
c.means = aggregate(age~exercise,data=newdata, mean)
print("(c) Group that has the lowest average age is is medium exercise group.")
d.sds = aggregate(age~exercise,data=newdata, sd)
print("(d) It appears the standard deviations of age by exercise level differ the most from the group L, low exercise level")

boxplot(sbp~educatn, data=newdata, main="Systolic Blood Pressure by Education",horizontal = TRUE)
print("(a) There is a trend that all groups are nearing each others")
library(ggplot2)
ggplot(newdata, aes(x=sbp))+geom_histogram(binwidth=2,color="black",fill="white")+facet_grid(married~.)+ggtitle("Systolic Blood Pressure by Marriage")
print("(b) Both group vary more than the other")
e_s=table(newdata$exercise,newdata$stress)
mosaicplot(e_s,main="Exercise vs Stress")
print("(c) The highest proportion of low stress is in the low exercise group")
m_s=table(newdata$married,newdata$stress)
mosaicplot(m_s,main="Marriage vs Stress")
print("(d) The highest proportion of low stress is in the married group")

table(newdata$exercise)
print("(a) Low exercise group had the most subject")
table(newdata$stress,newdata$exercise)
print("(b) Medium stress group had the most highly educated subjects")
s_a = aggregate(age~stress,data=newdata, mean)
s_a
print("(c) Medium stress group had the highest average age")
g_a = aggregate(sbp~gender,data=newdata, mean)
g_a
print("(d) Female had the lowest average systolic blood pressure")

result=t.test(sbp~married,data=newdata,var.equal=TRUE,level=0.95)
ts=round(result$statistic,4)
pval=result$p.value
CI=round(result$conf.int,4)
ts
print("(a) -1.3554")
pval
print("(b) 0.1758972")
CI
print("(c) (-8.3146  1.5259)")
print("(d) The confidence interval includes 0 meaning there are no significant difference between the groups")
print("(e) Since p value is larger than alpha null hypothesis is accepted and meaning that there are no relationship between two groups")
```
