---
title: "CHOE.JONGWOOK-HW2"
author: "JONG-WOOK-CHOE"
date: "4/16/2021"
output:
  pdf_document: default
  html_document: default
---

I.

```{r, echo=FALSE}
the.data=read.table("/Users/Bosco/Desktop/SPRING2021/STA 135/data/Smoke-Age-Death.txt",header = TRUE)
pairs(X20132 ~ X0 + X1 + X0.1, data = the.data)
```
(a) Side-by-side scatter plots (scatter plot matrix) created by pairs() function shows there are no linear relationship between variables.

(b) Since there are no linear relationship the data sets have no trend.

(c) Since there are no linear relationship there are no interaction between smoking status and death rates. 

II.

```{r, echo=FALSE}
the.data=read.table("/Users/Bosco/Desktop/SPRING2021/STA 135/data/Union_Army_Strength.txt",header = TRUE)
pairs(X1 ~ X1.1 + X708 + X42 + X262 +X83 + X321, data = the.data)
```
side-by-side scatter plots (scatter plot matrix) created by pairs() function shows there are some linear relationship between variables.

III.

```{r}
M1=matrix(c(1,0.5,0.5,1.25),ncol= 2,nrow =2)
M2=matrix(c(1,0.5,0.5,0.26),ncol= 2,nrow =2)
M3=matrix(c(1,0.5,0.5,0.01),ncol= 2,nrow =2)
inv1=solve(M1)
inv2=solve(M2)
inv3=solve(M3)
```

(a)

$$
M^{-1}=\frac{1}{1*1.25-0.5*0.5}\left(\begin{array}{cc}
1.25 & -0.5\\
-0.5 & 1
\end{array}\right)=
\left(\begin{array}{cc}
1.25 & -0.5\\
-0.5 & 1.0
\end{array}\right)
$$
(b)

$$
M^{-1}=\frac{1}{1*0.26-0.5*0.5}\left(\begin{array}{cc}
0.26 & -0.5\\
-0.5 & 1
\end{array}\right)=
\left(\begin{array}{cc}
26 & -50\\
-50 & 100
\end{array}\right)
$$
(c)

$$
M^{-1}=\frac{1}{1*0.01-0.5*0.5}\left(\begin{array}{cc}
0.01 & -0.5\\
-0.5 & 1
\end{array}\right)=
\left(\begin{array}{cc}
-0.42 & 2.08\\
2.08 & -4.17
\end{array}\right)
$$

IV.

```{r, echo=FALSE}
M=matrix(c(1,0.5,0.5,0.25),ncol= 2,nrow =2)
eigen(M)
t(eigen(M)$vectors)%*%M%*%eigen(M)$vectors
```

(a) Since the determinant is 0, it means not invertible. Also, the matrix is symmetric and $v^TMv\ge0$ for all $v\in V$. Therefore the matrix is semidefinite.

(b) Since it is not square matrix, it is not invertible. Also, the matrix is not symmetric therefore it is not positive definite or positive semidefinite.

V.

```{r}
round(M1%*%inv1) == diag(x=1,ncol=2,nrow=2)
round(M2%*%inv2) == diag(x=1,ncol=2,nrow=2)
round(M3%*%inv3) == diag(x=1,ncol=2,nrow=2)
```

(a) Since matrix multiplication $M*M^{-1}=I$, the matrix times its inverse results in the identity matrix.


```{r}
eig1=eigen(M1)$values
eig2=eigen(M2)$values
eig3=eigen(M3)$values
```

(b) Eigenvalues are 1.64 & 0.61 for first matrix, 1.252 & 0.007 for second matrix, and 1.209 & -0.199 for third matrix.

(c) To verify the eigen vectors they have to be satisfy $A\vec{x}=\lambda\vec{x}$. 

```{r}
eigv1=eigen(M1)$vectors
eigv2=eigen(M2)$vectors
eigv3=eigen(M3)$vectors
round(M1%*%eigv1[,1] - eig1[1]*eigv1[,1])
round(M1%*%eigv1[,2] - eig1[2]*eigv1[,2])
round(M2%*%eigv2[,1] - eig2[1]*eigv2[,1])
round(M2%*%eigv2[,2] - eig2[2]*eigv2[,2])
round(M3%*%eigv3[,1] - eig3[1]*eigv3[,1])
round(M3%*%eigv3[,2] - eig3[2]*eigv3[,2])
```

$A\vec{x}=\lambda\vec{x}$ therefore the eigen vectors satisfy corresponding equality.

```{r, echo=FALSE}
vl1=length(eigv1)
vl2=length(eigv2)
vl3=length(eigv3)
```

(d) Length of first vector is `r vl1`, second vector is `rvl2`, and third vector is `rvl3`.

(e)
```{r, echo=FALSE}
eigv1[1,]%*%eigv1[2,]
eigv2[1,]%*%eigv2[2,]
eigv3[1,]%*%eigv3[2,]

par(mfrow = c(1, 3))
par(pty = "s")
b1<-c(-1,1)
b2<-c(-1,1)
plot(x = b1, y = b2, type = "n", main =
expression(paste("Eigenvectors of ", M1)),
xlab = expression(b[1]), ylab = expression(b[2]) ,
panel.first=grid(col="gray", lty="dotted"))
abline(h = 0, lty = "solid", lwd = 2)
abline(v = 0, lty = "solid", lwd = 2)
arrows(x0 = 0, y0 = 0, x1 = eigv1[1,1], y1 =
eigv1[2,1], col = "red", lty = "solid")
arrows(x0 = 0, y0 = 0, x1 = eigv1[1,2], y1 =
eigv1[2,2], col = "red", lty = "solid")

par(pty = "s")
b1<-c(-1,1)
b2<-c(-1,1)
plot(x = b1, y = b2, type = "n", main =
expression(paste("Eigenvectors of ", M2)),
xlab = expression(b[1]), ylab = expression(b[2]) ,
panel.first=grid(col="gray", lty="dotted"))
abline(h = 0, lty = "solid", lwd = 2)
abline(v = 0, lty = "solid", lwd = 2)
arrows(x0 = 0, y0 = 0, x1 = eigv2[1,1], y1 =
eigv2[2,1], col = "red", lty = "solid")
arrows(x0 = 0, y0 = 0, x1 = eigv2[1,2], y1 =
eigv2[2,2], col = "red", lty = "solid")

par(pty = "s")
b1<-c(-1,1)
b2<-c(-1,1)
plot(x = b1, y = b2, type = "n", main =
expression(paste("Eigenvectors of ", M3)),
xlab = expression(b[1]), ylab = expression(b[2]) ,
panel.first=grid(col="gray", lty="dotted"))
abline(h = 0, lty = "solid", lwd = 2)
abline(v = 0, lty = "solid", lwd = 2)
arrows(x0 = 0, y0 = 0, x1 = eigv3[1,1], y1 =
eigv3[2,1], col = "red", lty = "solid")
arrows(x0 = 0, y0 = 0, x1 = eigv3[1,2], y1 =
eigv3[2,2], col = "red", lty = "solid")
```

The eigen vectors shown in graph are perpendicular to each other. Also, the product of the vectors is 0.

```{r}
det(M1)==eig1[1]*eig1[2]
round(det(M2))==round(eig2[1]*eig2[2])
round(det(M3))==round(eig3[1]*eig3[2])
```

(f) R calculation shows that the determinant of the matrix is equal to the product of the eigen values. 

```{r}
sum(diag(M1))==sum(eig1)
sum(diag(M2))==sum(eig2)
sum(diag(M3))==sum(eig3)
```

(g) R calculation shows that the trace of the matrix is equal to the sum of the eigen values.

(h) The matrix has to be symmetric, all eigen values are positive, and all the subdeterminants are also positive. Therefore first and second matrices are positive definite when third matrix is not because not all eigen values are positive. 


## Appendix

```{r,eval=FALSE}
the.data=read.table("/Users/Bosco/Desktop/SPRING2021/STA 135/data/Smoke-Age-Death.txt",header = TRUE)
pairs(X20132 ~ X0 + X1 + X0.1, data = the.data)
the.data=read.table("/Users/Bosco/Desktop/SPRING2021/STA 135/data/Union_Army_Strength.txt",header = TRUE)
pairs(X1 ~ X1.1 + X708 + X42 + X262 +X83 + X321, data = the.data)
M1=matrix(c(1,0.5,0.5,1.25),ncol= 2,nrow =2)
M2=matrix(c(1,0.5,0.5,0.26),ncol= 2,nrow =2)
M3=matrix(c(1,0.5,0.5,0.01),ncol= 2,nrow =2)
inv1=solve(M1)
inv2=solve(M2)
inv3=solve(M3)
round(M1%*%inv1) == diag(x=1,ncol=2,nrow=2)
round(M2%*%inv2) == diag(x=1,ncol=2,nrow=2)
round(M3%*%inv3) == diag(x=1,ncol=2,nrow=2)
eig1=eigen(M1)$values
eig2=eigen(M2)$values
eig3=eigen(M3)$values
eigv1=eigen(M1)$vectors
eigv2=eigen(M2)$vectors
eigv3=eigen(M3)$vectors
M1%*%eigv1[,1] == eig1[1]*eigv1[,1]
M1%*%eigv1[,2] == eig1[2]*eigv1[,2]
M2%*%eigv2[,1] == eig2[1]*eigv2[,1]
M2%*%eigv2[,2] == eig2[2]*eigv2[,2]
M3%*%eigv3[,1] == eig3[1]*eigv3[,1]
M3%*%eigv3[,2] == eig3[2]*eigv3[,2]
vl1=length(eigv1)
vl2=length(eigv2)
vl3=length(eigv3)
par(mfrow = c(1, 3))
par(pty = "s")
b1<-c(-1,1)
b2<-c(-1,1)
plot(x = b1, y = b2, type = "n", main =
expression(paste("Eigenvectors of ", M1)),
xlab = expression(b[1]), ylab = expression(b[2]) ,
panel.first=grid(col="gray", lty="dotted"))
abline(h = 0, lty = "solid", lwd = 2)
abline(v = 0, lty = "solid", lwd = 2)
arrows(x0 = 0, y0 = 0, x1 = eigv1[1,1], y1 =
eigv1[2,1], col = "red", lty = "solid")
arrows(x0 = 0, y0 = 0, x1 = eigv1[1,2], y1 =
eigv1[2,2], col = "red", lty = "solid")
eigv1[1,]%*%eigv1[2,]

par(pty = "s")
b1<-c(-1,1)
b2<-c(-1,1)
plot(x = b1, y = b2, type = "n", main =
expression(paste("Eigenvectors of ", M2)),
xlab = expression(b[1]), ylab = expression(b[2]) ,
panel.first=grid(col="gray", lty="dotted"))
abline(h = 0, lty = "solid", lwd = 2)
abline(v = 0, lty = "solid", lwd = 2)
arrows(x0 = 0, y0 = 0, x1 = eigv2[1,1], y1 =
eigv2[2,1], col = "red", lty = "solid")
arrows(x0 = 0, y0 = 0, x1 = eigv2[1,2], y1 =
eigv2[2,2], col = "red", lty = "solid")
eigv2[1,]%*%eigv2[2,]

par(pty = "s")
b1<-c(-1,1)
b2<-c(-1,1)
plot(x = b1, y = b2, type = "n", main =
expression(paste("Eigenvectors of ", M3)),
xlab = expression(b[1]), ylab = expression(b[2]) ,
panel.first=grid(col="gray", lty="dotted"))
abline(h = 0, lty = "solid", lwd = 2)
abline(v = 0, lty = "solid", lwd = 2)
arrows(x0 = 0, y0 = 0, x1 = eigv3[1,1], y1 =
eigv3[2,1], col = "red", lty = "solid")
arrows(x0 = 0, y0 = 0, x1 = eigv3[1,2], y1 =
eigv3[2,2], col = "red", lty = "solid")
eigv3[1,]%*%eigv3[2,]
det(M1)==eig1[1]*eig1[2]
round(det(M2))==round(eig2[1]*eig2[2])
round(det(M3))==round(eig3[1]*eig3[2])
sum(diag(M1))==sum(eig1)
sum(diag(M2))==sum(eig2)
sum(diag(M3))==sum(eig3)
```