---
title: "CHOE.JONGWOOK-HW5"
author: "JONG-WOOK-CHOE"
output: pdf_document
---

I. 

```{r, echo=FALSE}
cereal=read.csv("/Users/Bosco/Desktop/SPRING2021/STA 135/data/cereal.csv")
qqnorm(cereal[,5],main="Sugar Q-Q Plot")
qqline(cereal[,5], col='red')
qqnorm(cereal[,6],main="Fat Q-Q Plot")
qqline(cereal[,6], col='red')
qqnorm(cereal[,7],main="Sodium Q-Q Plot")
qqline(cereal[,7], col='red')
```

(a) Each of Q-Q plot shows marginal normality of the variables. However, it seems there might be outliers; notice two points in the upper-right and lower-left corner of the Q-Q plots.

```{r, echo=FALSE}
cerealdata=cereal[,5:7]
Xn<-colMeans(cerealdata)
Sn<-var(cerealdata)
Sninv<-solve(Sn)
datachisq<-diag(t(t(cerealdata)-Xn)%*%Sninv%*%(t(cerealdata)-Xn))
qqplot(qchisq(ppoints(500), df = 3), datachisq,
main = expression("Chi-square plot for" ~~ {chi^2}[nu == 3]))
qqline(datachisq,distribution=function(p) qchisq(p, df = 3))
```

For the most part the chi-square plot is skewed to the left but it shows a linear pattern. There is a point, upper right corner, which shows deviation from the linear trend. It may indicate that it is a outlier.

```{r, echo=FALSE}
library(RVAideMemoire)
mqqnorm(cerealdata,main="Q-Q plot of Mahalanobis D^2 VS. quantiles of Chi^2")
```

(b) The chi-square Q-Q plot shows a linear pattern. Therefore it indicated it follows multivariate normal distribution.

"A popular way to identify and deal with multivariate outliers is to use Mahalanobis Distance. It calculates the distance of each case from the central mean. Larger values indicate that a case is farther from where most of the points cluster."

Therefore, #26 and #25 are the two outliers from the plot.

II.

```{r, echo=FALSE}
p<-3
N<-40
mu.hat<-colMeans(cerealdata)
sigma.hat<-cov(cerealdata)
mu.Ho<-c(9.8,1.7,156.8)
T.sq<-N*t(mu.hat-mu.Ho)%*%solve(sigma.hat)%*%(mu.hat-mu.Ho)
test.stat<-(N-p)/(p*(N-1))*T.sq
crit.val<-qf(0.95, p, N-p)
p.value<-1-pf((N-p)/(p*(N-1))*T.sq, p, N-p)
round(data.frame(T.sq, test.stat, crit.val, p.value), 3)
```

(a) $T^2$ for testing the null hypothesis $H_0:\mu=[9.8,1.7,156.8]^T$ is 19.742 where $\mu_1,\mu_2,$and$\mu_3$ represent the population mean of sugar, fat, and sodium cereal contain.

(b) $E(t)=0$
    $Var(t)=\frac{n-1}{n-3}$
    $T^2$~[0,1.05]

(c) Using the results in the previous parts, $H_0$ at the level $\alpha=0.01$ is smaller than the p-value. Therefore we fail to reject the null at the $\alpha$ level and conclude that the population true vector mean is $\mu=\mu_0=[9.8,1.7,156.8]^T$

(d) With more reasonable $H_0$ at the level $\alpha=0.05$ p-value is smaller than the critical level therefore we reject our null hypothesis and conclude that $\mu\not=[9.8,1.7,156.8]^T$

III.

```{r, echo=FALSE}
sweat=read.table("/Users/Bosco/Desktop/SPRING2021/STA 135/data/Sweet_data.txt")
colnames(sweat)=c("sweet_rate","sodium","potassium")
p<-3
N<-20
mu.hat<-colMeans(sweat)
sigma.hat<-var(sweat)
mu.Ho<-c(3.5,45,15)
T.sq<-N*t(mu.hat-mu.Ho)%*%solve(sigma.hat)%*%(mu.hat-mu.Ho)
test.stat<-(N-p)/(p*(N-1))*T.sq
crit.val<-qf(0.95, p, N-p)
p.value<-1-pf((N-p)/(p*(N-1))*T.sq, p, N-p)
round(data.frame(T.sq, test.stat, crit.val, p.value), 5)
```

Since the p-value is 1.755542e-08, it is too small at any reasonable $\alpha$ and we reject our null hypothesis. Therefore conclude that the true population vector mean of the sweat data $\mu$ is not the hypothesized mean $\mu\not=[3.5,45,15]^T$

IV.

```{r, echo=FALSE}
BF.Lowermu1<-mu.hat[1]-qt(1-(0.05/2),N-1)*sqrt(sigma.hat[1,1]/N)
BF.Uppermu1<-mu.hat[1]+qt(1-(0.05/2),N-1)*sqrt(sigma.hat[1,1]/N)
BF.Lowermu2<-mu.hat[2]-qt(1-(0.05/2),N-1)*sqrt(sigma.hat[2,2]/N)
BF.Uppermu2<-mu.hat[2]+qt(1-(0.05/2),N-1)*sqrt(sigma.hat[2,2]/N)
BF.Lowermu3<-mu.hat[3]-qt(1-(0.05/2),N-1)*sqrt(sigma.hat[3,3]/N)
BF.Uppermu3<-mu.hat[3]+qt(1-(0.05/2),N-1)*sqrt(sigma.hat[3,3]/N)
BF.CI4mu1<-c(BF.Lowermu1,BF.Uppermu1)
BF.CI4mu2<-c(BF.Lowermu2,BF.Uppermu2)
BF.CI4mu3<-c(BF.Lowermu3,BF.Uppermu3)
round(data.frame(BF.CI4mu1, BF.CI4mu2, BF.CI4mu3), 2)
```
On the cereal data testing equality of mean sugar, fat, and sodium, content across different shelf using Berforoni's method tells the simultaneous confidence intervals for $\mu_1$, $\mu_2$, and $\mu_3$. Therefore with 95% confident each $\mu$ stays between (3.85 5.43), (38.78 52.02), and (9.07 10.86) using the bonferroni correction.

## Appendix

```{r,eval=FALSE}
cereal=read.csv("/Users/Bosco/Desktop/SPRING2021/STA 135/data/cereal.csv")
qqnorm(cereal[,5],main="Sugar Q-Q Plot")
qqline(cereal[,5], col='red')
qqnorm(cereal[,6],main="Fat Q-Q Plot")
qqline(cereal[,6], col='red')
qqnorm(cereal[,7],main="Sodium Q-Q Plot")
qqline(cereal[,7], col='red')
cerealdata=cereal[,5:7]
Xn<-colMeans(cerealdata)
Sn<-var(cerealdata)
Sninv<-solve(Sn)
datachisq<-diag(t(t(cerealdata)-Xn)%*%Sninv%*%(t(cerealdata)-Xn))
qqplot(qchisq(ppoints(500), df = 3), datachisq,
main = expression("Chi-square plot for" ~~ {chi^2}[nu == 3]))
qqline(datachisq,distribution=function(p) qchisq(p, df = 3))
library(RVAideMemoire)
mqqnorm(cerealdata,main="Q-Q plot of Mahalanobis D^2 VS. quantiles of Chi^2")
p<-3
N<-40
mu.hat<-colMeans(cerealdata)
sigma.hat<-cov(cerealdata)
mu.Ho<-c(9.8,1.7,156.8)
T.sq<-N*t(mu.hat-mu.Ho)%*%solve(sigma.hat)%*%(mu.hat-mu.Ho)
test.stat<-(N-p)/(p*(N-1))*T.sq
crit.val<-qf(0.95, p, N-p)
p.value<-1-pf((N-p)/(p*(N-1))*T.sq, p, N-p)
round(data.frame(T.sq, test.stat, crit.val, p.value), 3)
sweat=read.table("/Users/Bosco/Desktop/SPRING2021/STA 135/data/Sweet_data.txt")
colnames(sweat)=c("sweet_rate","sodium","potassium")
p<-3
N<-20
mu.hat<-colMeans(sweat)
sigma.hat<-var(sweat)
mu.Ho<-c(3.5,45,15)
T.sq<-N*t(mu.hat-mu.Ho)%*%solve(sigma.hat)%*%(mu.hat-mu.Ho)
test.stat<-(N-p)/(p*(N-1))*T.sq
crit.val<-qf(0.95, p, N-p)
p.value<-1-pf((N-p)/(p*(N-1))*T.sq, p, N-p)
round(data.frame(T.sq, test.stat, crit.val, p.value), 5)
BF.Lowermu1<-mu.hat[1]-qt(1-(0.05/2),n-1)*sqrt(sigma.hat[1,1]/n)
BF.Uppermu1<-mu.hat[1]+qt(1-(0.05/2),n-1)*sqrt(sigma.hat[1,1]/n)
BF.Lowermu2<-mu.hat[2]-qt(1-(0.05/2),n-1)*sqrt(sigma.hat[2,2]/n)
BF.Uppermu2<-mu.hat[2]+qt(1-(0.05/2),n-1)*sqrt(sigma.hat[2,2]/n)
BF.Lowermu3<-mu.hat[3]-qt(1-(0.05/2),n-1)*sqrt(sigma.hat[3,3]/n)
BF.Uppermu3<-mu.hat[3]+qt(1-(0.05/2),n-1)*sqrt(sigma.hat[3,3]/n)
BF.CI4mu1<-c(BF.Lowermu1,BF.Uppermu1)
BF.CI4mu2<-c(BF.Lowermu2,BF.Uppermu2)
BF.CI4mu3<-c(BF.Lowermu3,BF.Uppermu3)
round(data.frame(BF.CI4mu1, BF.CI4mu2, BF.CI4mu3), 2)
```